---
title: "Early, Simple Models"
output: html_notebook
---

### Model 1- Single level model w/o Priors, Generated Quantities

We'll just model price as a function of UPC. 

```{r}
writeLines(readLines("models/1_single_level_no_prior_no_gen.stan"))
```

Next, we'll compile the model. This allows us to check for errors.

```{r}
m1_upc <- stanc(file = "models/1_single_level_no_prior_no_gen.stan") # Check Stan file
m1_upc <- stan_model(stanc_ret = m1_upc) # Compile Stan code
```

When running 4 chains in parallel, the time to execute is as long as the longest chain. 

```{r}
fit1_upc <- sampling(m1_upc, #use sampling function, rather than stan()
                warmup=1000, #recommended in Betancourt "RStan Workflow"
                iter=2000, #TOTAL iterations- first 1000 are warmup
                seed=1, #ensures reproducibility
                data=data, 
                chains = 4, #each chain has different initial condition which helps to find pathological neighborhood of posterior
                cores = getOption("mc.cores", 1L),
                verbose = TRUE)
rm(m1_upc)
```

```{r}
pairs(fit1_upc)
```


```{r}
list_of_draws <- rstan::extract(fit1_upc)
print(names(list_of_draws))
print(fit1_upc, c("beta_0", "beta_1")) 
```

### Model 2- Single level model w/ Priors, w/o Generated Quantities

We'll add some priors. 

```{r}
writeLines(readLines("models/2_single_level_priors_no_gen.stan"))
```

Next, we'll compile the model. This allows us to check for errors.

```{r}
m2_upc <- stanc(file = "models/2_single_level_priors_no_gen.stan") # Check Stan file
m2_upc <- stan_model(stanc_ret = m2_upc) # Compile Stan code
```

67 seconds total.

```{r}
fit2_upc <- sampling(m2_upc, 
                warmup=400, 
                iter=4000, 
                seed=1, 
                data=data, 
                chains = 1,
                cores = getOption("mc.cores", 1L),
                verbose = TRUE)
rm(m2_upc)
```

```{r}
list_of_draws <- rstan::extract(fit1_upc)
print(names(list_of_draws))
print(fit1_upc, c("beta_0", "beta_1")) 
```
